### 2019-12-25 paper/paper.tex
fixed wrong opam file

\documentclass[a4paper, 11pt]{article}

%% Language and font encodings. This says how to do hyphenation on end of lines.
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage{listingsutf8}
\usepackage{xcolor}
\usepackage{multicol}
\setlength{\columnsep}{1cm}
\setlength{\columnseprule}{0.5pt}
%\usepackage{aas_macros}

%% Sets page size and margins. You can edit this to your liking
\usepackage[top=1.3cm, bottom=2.0cm, outer=2.5cm, inner=2.5cm, heightrounded,
marginparwidth=1cm, marginparsep=0.4cm, margin=1.5cm]{geometry}

%% Useful packages
\usepackage{graphicx} %allows you to use jpg or png images. PDF is still recommended
\usepackage[colorlinks=False]{hyperref} % add links inside PDF files
\usepackage{amsmath}  % Math fonts
\usepackage{amsfonts} %
\usepackage{amssymb}  %

\definecolor{vgreen}{RGB}{104,180,104}
\definecolor{vblue}{RGB}{49,49,255}
\definecolor{vorange}{RGB}{255,143,102}
\lstdefinestyle{bash} {language=bash, basicstyle=\ttfamily,
	keywordstyle=\color{vblue}, identifierstyle=\color{black},
	commentstyle=\color{vgreen}, tabsize=4,
%	moredelim=*[s][\colorIndex]{[}{]},
	literate=*{:}{:}1}

\lstdefinestyle{caml} {language=caml, basicstyle=\ttfamily, columns=[c]fixed,
 keywordstyle=\color{vblue}, identifierstyle=\color{black},
 commentstyle=\color{vgreen}, upquote=true, commentstyle=, breaklines=true,
 showstringspaces=false, stringstyle=\color{blue},
 literate={'"'}{\textquotesingle "\textquotesingle}3}

%% Citation package
\usepackage[authoryear]{natbib}
\bibliographystyle{abbrvnat}
\setcitestyle{authoryear,open={(},close={)}}

\title{Minicaml, a purely functional, didactical programming language with an
interactive REPL. WORK IN PROGRESS DRAFT}
\author{Alessandro Cheli\\Course taught by Prof. Gianluigi Ferrari\\and Prof. Francesca Levi}

\begin{document}
\maketitle

\begin{abstract}
\textbf{minicaml} is a small, purely functional interpreted programming language
with a didactical purpose. It is based on the Prof. Gianluigi Ferrari and Prof.
Francesca Levi's minicaml, an evaluation example to show students attending the
Programming 2 course at the University of Pisa how interpreters work. It is an
interpreted language featuring a Caml-like syntax, static (lexical scoping),
interchangeable eager and lazy evaluation and a didactical REPL that
shows each AST expression and each evaluation step.
\end{abstract}

\section{REPL and command line interface}
\subsection{Installation}
\textbf{minicaml} is available in the opam 2.0 repository.
(\url{https://opam.ocaml.org/}). The easiest way to install minicaml is with the
OCaml package manager \textbf{opam}. To do so, please check that you have a version of opam $\geq$
2.0.0 and run:
\begin{lstlisting}[style=bash]
opam install minicaml
\end{lstlisting}
Alternatively, \textbf{minicaml} can be installed from source by downloading the
source code git repository and building it manually. \textbf{minicaml} has been tested
only on Linux and macOS systems. It has not been tested yet on Windows and BSD
derived systems.
\begin{lstlisting}[style=bash]
# download the source code
git clone https://github.com/0x0f0f0f/minicaml
# cd into the source code directory
cd minicaml
# install dependencies
opam install ANSITerminal dune ppx_deriving menhir cmdliner
# compile
make
# execute
make run
# install
make install
\end{lstlisting}

\begin{multicols}{2}
\section{Lexer}

\section{Parser}

\section{AST Optimization}
Before being evaluated, AST expressions are analyzed and optimized by an
optimizer function that is recursively called over the tree that is representing
the expression. The optimizer simplifies expressions which result is known and
therefore does not need to be evaluated. For example, it is known that \texttt{5
+ 3 $\equiv$ 8} and \texttt{true \&\& (true || (false \&\& false)) $\equiv$
true}. When a programmer writes a program, she or he may not want to do all the
simple calculations before writing the program in which they appear in, we rely
on machines to simplify those processes. Reducing constants before evalaution
may seem unnecessary when writing a small program, but they do take away
computation time, and if they appear inside of loops, it is a wise choice to
simplify those constant expressions whose result is already known before it is
calculated in all the loop iterations. It is also necessary in optimizing
programs before compilation. The optimizer, by now, reduces operations between
constants and \texttt{if} statements whose guard is always true (or false). To
achieve minimization to an unreducible form, optimizer calls are repeated until
it produces an output equal to its input; this way, we get a tree representing
an expression that cannot be optimized again. This process is fairly easy:

\begin{lstlisting}[style=caml]
let rec iterate_optimizer e =
	let oe = optimize e in
	if oe = e then e (* Bottoms out *)
	else iterate_optimizer oe
\end{lstlisting}

\section{Types}

\section{Evaluation}

\section{Tests}
Unit testing is extensively performed using the alcotest testing framework. Code
coverage is provided by the \texttt{bisect\_ppx} library which yields an HTML
page containing the coverage percentage when unit tests are run by the dune
build system. After each commit is pushed to the remote version control repository on
Github, the package is built and tests are run thanks to the Travis Continuos
Integration system.

\section{Thanks to}

\begin{itemize}
	\item Prof. Gian-Luigi Ferrari for teaching us how to project and develop
	interpreters in OCaml
	\item Antonio DeLucreziis for helping me implement lazy evaluation.
	\item Prof. Alessandro Berarducci for helping me study lambda calculus in deep.
\end{itemize}

\end{multicols}
\end{document}